---
title: "Barchart fundamentals, draw gender"
author: "Steve Simon"
date: "2019-08-03"
output: powerpoint_presentation
---

```{r echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("../../common-files/setup.R")
suppressMessages(suppressWarnings(library(reticulate)))
titanic <- read_tsv("../data/titanic.txt")
```

### Barchart fundamentals, Gender barchart
+ Draw a bar chart showing the number of people in each gender (use the Sex variable).

### Barchart fundamentals, Gender barchart, Python code

((Python code))

### Barchart fundamentals, Gender barchart, Python results

```{python}
print("Hello world")
a = 3
print(a + 2)
```

((Python results))

### Barchart fundamentals, Gender barchart, R code

```{}
ggplot(titanic, aes(Pclass)) +
  geom_bar()
```

### Barchart fundamentals, Gender barchart, R results

```{r bender-barchart}
ggplot(titanic, aes(Sex)) +
  geom_bar()
```

### Barchart fundamentals, Gender barchart, Tableau steps

((Tableau steps))

### Barchart fundamentals, Gender barchart, Tableau results

((Tableau results))

