---
title: "Exercise, change line defaults"
author: "Steve Simon"
date: "Created: 2019-08-19"
output: powerpoint_presentation
---

```{r child="../../common-files/src/component-header.Rmd"}
```

```{r read-files}
load("../../common-files/data/cpi.RData")
load("../../common-files/data/houses.RData")
```

### Exercise, change defaults
+ Draw a line graph with the cpi data
  + x=t
  + y=CPI
+ Change the defaults for the line
  + Use a dashed rather than a solid lineline to a dashed line.
  + Make the width equal to 3
  + Make the color green

<div class="notes">

Take the CPI line graph and modify some of the default options. Make the line green and increase the thickness to three pixels. Use a dashed line instead of a dotted line.

</div>

### Exercise, Python code
+ Here's the Python code
```{python}
import pandas as pd
import altair as alt
df = pd.read_csv("../../common-files/data/cpi.csv")
```

```{python, echo=TRUE}
ch = alt.Chart(df).mark_line(
    strokeDash=[5, 5, 2, 5]
).encode(
    alt.X('t'),
    alt.Y('CPI', scale=alt.Scale(domain=(200, 400)))
)
```

```{python}
ch.save("../images/python/dotted-line.html")
```

### Exercise, Python output

### Exercise, R code
+ Here's the R code
```
ggplot(cpi, aes(x=t, y=CPI)) +
  geom_line(size=3, linetype="dashed", color="green")
```

### Exercise, R output

```{r width-shape-color}
initiate_image()
ggplot(cpi, aes(x=t, y=CPI)) +
  geom_line(size=3, linetype="dashed", color="green")
finalize_image()
```

`r display_image`

<div class="notes">



</div>


### Exercise, Tableau steps

### Exercise, Tableau output


