---
title: "Recommendations for line graphss"
author: "Steve Simon"
date: "Created: 2019-08-19"
output: powerpoint_presentation
---

```{r child="../../common-files/src/component-header.Rmd"}
```

### Recomendations, line graphs
+ High frequenct events
  + Avoid think lines, long dashes.
  
### Recommendation, Avoid thick dashed lines for high frequency data

```{r thick-versus-thin}
load("../../common-files/data/cpi.RData")
initiate_image()
ggplot(cpi, aes(t, CPI)) +
  geom_line(size=3, linetype="33", color="black") +
  geom_line(size=1, linetype="solid", color="red")
finalize_image()
```

`r display_image`

<div class="notes">

High frequency events are sudden surges and dips in a line over a narrow range. If your data has high frequency events (think stock market prices), then avoid thick lines or long dashes. These can hide some of the high frequency events.

</div>

### Recommendations, Avoid error bars (1/3)

```{r average-with-standard-deviation}
load("../../common-files/data/houses.RData")
saratoga_houses %>%
  group_by(Bedrooms) %>%
  summarize(avg_price=mean(Price), sd_price=sd(Price), n=n()) -> aggregated_prices
initiate_image()
ggplot(aggregated_prices, aes(Bedrooms, avg_price)) +
  geom_line() +
  geom_errorbar(aes(ymin=avg_price-sd_price, ymax=avg_price+sd_price), width=0.1)
finalize_image("Average price shown with error bar equal to one standard deviaiton")
```

`r display_image`

<div class="notes">

There is a clear and consistent trend in the average price. As the number of bedrooms increase, the average price increases.

</div>

### R output

```{r average-with-standard-error}
load("../../common-files/data/houses.RData")
saratoga_houses %>%
  group_by(Bedrooms) %>%
  summarize(avg_price=mean(Price), sd_price=sd(Price), n=n()) -> aggregated_prices
initiate_image()
ggplot(aggregated_prices, aes(Bedrooms, avg_price)) +
  geom_line() +
  geom_errorbar(aes(ymin=avg_price-sd_price/sqrt(n), ymax=avg_price+sd_price/sqrt(n)), width=0.1)
finalize_image("Average price shown with error bar equal to one standard error")
```

`r display_image`

<div class="notes">

There is a clear and consistent trend in the average price. As the number of bedrooms increase, the average price increases.

</div>

### Recommendations, show variation
+ Original data
+ Error bars
+ Confidence bands
+ Boxplots

### Recommendations, when to start at zero

### Recommendations, differentiating lines

### Recommendations, lines imply continuity, bars don't

### Recommendations, aspect ratio

### Recommendations, square graphs

### Recommendations, sort your data or hope that your software does