---
title: "Data-viz=03, Tutorial on Gestalt principles"
author: "Steve Simon"
date: "Created: 2019-08-19"
output: powerpoint_presentation
---

```{r echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("../../common-files/setup.R")
```

### More theory -- The gestalt of graphics

+ These ideas drawn from the Bergen and Iverson workshop.
+ Gestalt definition
  + "The whole is greater than the sum of the parts"
+ How do you draw someone's eye to quickly make certain associations?
  
<div class="notes">



</div>

### Gestalt theory in artistic design
+ Many lesssons in effective artistic design
  + [The Gestalt Principles](https://graphicdesign.sfcc.spokane.edu/dZine/tutorials/process/gestaltprinciples/gestaltprinc.htm) Spokane Falls Community College. 
  + [Gestalt Theory](https://www.sophia.org/tutorials/gestalt-theory-2) Sophia.
  + [Gestalt Principles Applied to Design](https://www.grayboxpdx.com/blog/post/gestalt-principles-applied-to-design) The Graybox blog, January 19, 2015.
  + [Gestalt Principles](https://www.interaction-design.org/literature/topics/gestalt-principles) Interaction Design Foundation.
  
<div class="notes">

The field of artistic design is more commonly called graphic design, but I use the former term to distinguish it from statistical graphics. Artistic design is the use of typography, photography, and illustration to effectively convey a message. Artistic designers work on commercial logos, magazines and brochure s, and product packaging.

</div>

### A block of points - no emphasis

```{r block-no-emphasis}
initiate_image()
df <- data.frame(
  x=rep(1:5, 4),
  y=rep(1:4, each=5)
)
ggplot(df, aes(x=x, y=y)) +
  scale_x_continuous(name=NULL, breaks=NULL) +
  scale_y_continuous(name=NULL, breaks=NULL) -> g
g + geom_point(size=5)
finalize_image()
```

`r display_image`

<div class="notes">

What you see in this graph is 20 individual points. When you design a graph, you want people to group things in a way to reveal patterns.

</div>

### Color emphasizes rows

```{r block-color}
initiate_image()
f <- rep(c("red", "green", "red", "green"), each=5)
g + geom_point(color=f, size=5)
finalize_image()
```

`r display_image`

<div class="notes">



</div>


### Shape emphasizes columns

```{r block-shape}
initiate_image()
f <- rep(c("triangle", "square", "triangle", "square", "triangle"), 4)
g + geom_point(shape=f, size=5)
finalize_image()
```

`r display_image`

<div class="notes">



</div>

### Double up for stronger emphasis

```{r block-double-up}
initiate_image()
f1 <- rep(c("red", "green", "red", "green"), each=5)
f2 <- rep(c("triangle", "square", "triangle", "square"), each=5)
g + geom_point(color=f1, shape=f2, size=5)
finalize_image()
```

`r display_image`

<div class="notes">



</div>

### Connectedness emphasizes rows

```{r block-connectedness}
initiate_image()
g + geom_point(size=5) +
  geom_segment(x=1, y=1, xend=5, yend=1) + 
  geom_segment(x=1, y=2, xend=5, yend=2) + 
  geom_segment(x=1, y=3, xend=5, yend=3) + 
  geom_segment(x=1, y=4, xend=5, yend=4) 
finalize_image()
```

`r display_image`

<div class="notes">



</div>

### Proximity emphasize columns

```{r block-proximity}
initiate_image()
g + geom_point(size=5) +
  expand_limits(y=c(-3, 8))
finalize_image()
```

`r display_image`

<div class="notes">



</div>

### Enclosure for emphasis, eight special points

```{r block-enclosure}
initiate_image()
g + geom_point(size=5) +
  expand_limits(y=c(0.8, 4.2)) +
  expand_limits(x=c(0.8, 5.2)) +
  geom_segment(x=3.8, y=0.8, xend=5.2, yend=0.8) +
  geom_segment(x=3.8, y=4.2, xend=5.2, yend=4.2) +
  geom_segment(x=3.8, y=0.8, xend=3.8, yend=4.2) +
  geom_segment(x=5.2, y=0.8, xend=5.2, yend=4.2)
finalize_image()
```

`r display_image`

<div class="notes">



</div>

### Continuity

((Show example of sloping text))

### Changing your emphasis

((Recreate example from slides 47-50 of Bergen and Iverson))

